---
title: "Homework 1: Amelia Grant-Alfieri, ag3911"
output: github_document
---

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



Before beginning, we must load the tidyverse package using `library(tidyverse)`.

```
library(tidyverse)
```

# Problem 1

First, we create a dataframe containing four elements:
- A random sample of size 10 from a uniform [0,5] distribution
- A logical vector indicating whether elements of the sample are greater than 2
- A (length-10) character vector
- A (length-10) factor vector. 
```{r}
p1_df = tibble(
  x = runif(10, min = 0, max = 5),
  vec_log_above2 = c( x > 2),
  vec_char = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten"),
  vec_factor = factor(c("male", "female", "male", "female", "female", "female", "female", "male", "male", "female"))
)
```


Second, we caluclate the mean of each variable. Then we analyze what works and what doesn't work and why.
```{r}
mean(vec_log_above2)
mean(vec_factor)
mean(vec_char)
```
This runs smoothly for the numeric vector 'vec_log_above2' and generates a mean = 0.6. However, it produces an error stating that 'vec_char' and 'vec_factor' are not found. They are not numeric variables, and we can't calculate a mean for a non-numeric variable. 


Third, we apply the as.numeric function to the logical, character, and factor variables. Then we describe what happens.
```{r}
p1_df$vec_log_above2 <- as.numeric(p1_df$vec_log_above2)
p1_df$vec_char <- as.numeric(p1_df$vec_char)
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For 'vec_log_above2', 'TRUE' is changed to 1 and 'FALSE' is changed to 0.
For 'vec_char', all values read 'NA' because characters cannot be read as categorical or numeric.
For 'vec_factor', 'male' is changed to 2 and 'female' is changed to 1.  


Fourth, we convert the character variable to factor to numeric and convert the factor variable from factor to character to numeric. Then we describe what happens. 
```{r}
p1_df$vec_char <- as.factor(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column label changes from <dbl> to <fct>
```{r}
p1_df$vec_char <- as.numeric(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column is relabeled <dbl>
```{r}
p1_df$vec_factor <- as.character(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column label changes from <dbl> to <chr>
```{r}
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column is relabled <dbl>
```


# Problem 2

First, we create a data frame containing five elements:
- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the  x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector

```{r}
p2_df = tibble(
  x = rnorm(1000)
  y = rnorm(1000)
  vec_log_sumabove0 = c( x + y > 0 )
  vec_num = 
  vec_factor = #labeling should be differene 
)
```

Second, we describe the vector using inline R code, including: i) the size of the dataset, ii) the mean and median of x, and iii) the proportion of cases for which the logical vector is  TRUE.

for histogram of x variable:
```{r x_hist}
ggplot(plot_df, aes(x = x)) + geom_histogram()
```

Third, we make a scatterplot of y vs x and color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). 

```{r yx_scatter}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

Fourth, we make a scatterplot that color points using the numeric variable. Then we comment on the color scale. 

```{r yx_scatter}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

Fifth, we make a scatterplot that color points using the factor variable. Then we comment on the color scale.

```{r yx_scatter}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

Last, we export our first logical variable scatterplot to the project directory using  ggsave.

```{r}
```




