---
title: "Homework 1: Amelia Grant-Alfieri, ag3911"
output: github_document
---


Before beginning, load the tidyverse package. 
```{r include=FALSE}
library(tidyverse)
```

# Problem 1

## Create a dataframe containing four elements:
* A random sample of size 10 from a uniform [0,5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector
```{r}
set.seed(1)
p1_df = tibble(
  x = runif(10, min = 0, max = 5),
  vec_log_above2 = c(x > 2),
  vec_char = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten"),
  vec_factor = factor(c("male", "female", "male", "female", "female", "female", "female", "male", "male", "female"))
)
```


## Calculate the mean of each variable. Analyze what works, what doesn't, and why.
```{r}
mean(p1_df$vec_log_above2)
mean(p1_df$vec_factor)
mean(p1_df$vec_char)
```
This runs smoothly for the numeric vector 'vec_log_above2' and generates a mean = 0.6. However, it produces an error stating that 'vec_char' and 'vec_factor' are not found. They are not numeric variables, and we can't calculate a mean for a non-numeric variable. 


## Apply the as.numeric function to the logical, character, and factor variables. Describe what happens.
```{r eval=FALSE}
p1_df$vec_log_above2 <- as.numeric(p1_df$vec_log_above2)
p1_df$vec_char <- as.numeric(p1_df$vec_char)
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For 'vec_log_above2', 'TRUE' is changed to 1 and 'FALSE' is changed to 0.
For 'vec_char', all values read 'NA' because characters cannot be read as categorical or numeric.
For 'vec_factor', 'male' is changed to 2 and 'female' is changed to 1.  


## Convert the character variable to factor and then to numeric. Convert the factor variable from factor to character and then to numeric. Describe what happens. 
```{r}
p1_df$vec_char <- as.factor(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column label changes from 'dbl' to 'fct'.
```{r}
p1_df$vec_char <- as.numeric(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column is relabeled 'dbl'.
```{r}
p1_df$vec_factor <- as.character(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column label changes from 'dbl' to 'chr'.
```{r}
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column is relabeled 'dbl'.



# Problem 2

## Create a data frame containing five elements:
* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the  x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector
```{r}
set.seed(1)
p2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_log_sumabove0 = c(x + y > 0),
  vec_num = as.numeric(vec_log_sumabove0),
  vec_factor = as.factor(vec_log_sumabove0),
  )
```

## Describe the vector. 
The dataset consists of 1,000 observations and 5 columns. The mean of x is `r mean(p2_df$x)`. The median of x is `r median(p2_df$x)`. ` table(p2_df$vec_num == 1)` determines that there are 490 TRUE cases for the logical vector. Dividing the number of true values by the total number of values: `490/1000`, we find that 49% is the proportion of cases for which the logical vector is TRUE.

## Make a scatterplot of y vs x and color points using the logical variable.
```{r}
splot1 <- ggplot(p2_df, aes(x = x, y = y, color = vec_log_sumabove0)) + geom_point()
splot1
```
The points are one of two colors and the coloring is split along a diagonal line. 

## Make a scatterplot that colors points using the numeric variable. Describe the color scale. 
```{r}
splot2 <- ggplot(p2_df, aes(x = x, y = y, color = vec_num)) + geom_point()
splot2
```
The points are colored on a gradient from dark to light and split along a diagonal line as in the previous plot. Light points are on the upper/right side and dark points are on the lower/left side of the diagonal line. 

## Make a scatterplot that colors points using the factor variable. Describe the color scale.
```{r}
splot3 <- ggplot(p2_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
splot3
```
The points are colored and distributed in the same manner as the first plot. 

## Export the first logical variable scatterplot to the project directory.
```{r}
ggsave("hw1_p2_splot1.png", plot = splot1, device = png())
```




