---
title: "Homework 1: Amelia Grant-Alfieri, ag3911"
output: github_document
---


Before beginning, we must load the tidyverse package using `library(tidyverse)`.

```{r include=FALSE}
library(tidyverse)
```

# Problem 1

First, we create a dataframe containing four elements:
- A random sample of size 10 from a uniform [0,5] distribution
- A logical vector indicating whether elements of the sample are greater than 2
- A (length-10) character vector
- A (length-10) factor vector. 
```{r}
set.seed(1)
p1_df = tibble(
  x = runif(10, min = 0, max = 5),
  vec_log_above2 = c(x > 2),
  vec_char = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten"),
  vec_factor = factor(c("male", "female", "male", "female", "female", "female", "female", "male", "male", "female"))
)
```


Second, we caluclate the mean of each variable. Then we analyze what works and what doesn't work and why.
```{r}
mean(p1_df$vec_log_above2)
mean(p1_df$vec_factor)
mean(p1_df$vec_char)
```
This runs smoothly for the numeric vector 'vec_log_above2' and generates a mean = 0.6. However, it produces an error stating that 'vec_char' and 'vec_factor' are not found. They are not numeric variables, and we can't calculate a mean for a non-numeric variable. 


Third, we apply the as.numeric function to the logical, character, and factor variables. Then we describe what happens.
```{r}
p1_df$vec_log_above2 <- as.numeric(p1_df$vec_log_above2)
p1_df$vec_char <- as.numeric(p1_df$vec_char)
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For 'vec_log_above2', 'TRUE' is changed to 1 and 'FALSE' is changed to 0.
For 'vec_char', all values read 'NA' because characters cannot be read as categorical or numeric.
For 'vec_factor', 'male' is changed to 2 and 'female' is changed to 1.  


Fourth, we convert the character variable to factor to numeric and convert the factor variable from factor to character to numeric. Then we describe what happens. 
```{r}
p1_df$vec_char <- as.factor(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column label changes from <dbl> to <fct>
```{r}
p1_df$vec_char <- as.numeric(p1_df$vec_char)
```
For vec_char, values remain 'NA' but the column is relabeled <dbl>
```{r}
p1_df$vec_factor <- as.character(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column label changes from <dbl> to <chr>
```{r}
p1_df$vec_factor <- as.numeric(p1_df$vec_factor)
```
For vec_factor, values remain 2s or 1s, but the column is relabled <dbl>
```


# Problem 2

First, we create a data frame containing five elements:
- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the  x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector
```{r}
set.seed(1)
p2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_log_sumabove0 = c(x + y > 0),
  vec_num = as.numeric(vec_log_sumabove0),
  vec_factor = as.factor(vec_log_sumabove0),
  )
```

Second, we describe the vector. We use `p2_df` to find the size of the dataset, which is 5,000 values made up of 1,000 rows and 5 columns. We use `r mean(p2_df$x)` to find the mean of x, which is -0.0245. 
We use `r median(p2_df$x)` to find the median of x, which is -0.0436.  
We use `r table(p2_df$vec_num == 1)` to determine that there are 490 TRUE cases for the logical vector. Then we calculate that `490/1000` or 0.49 is the proportion of cases for which the logical vector is TRUE.


Third, we make a scatterplot of y vs x and color points using the logical variable.
```{r}
splot1 <- ggplot(p2_df, aes(x = x, y = y, color = vec_log_sumabove0)) + geom_point()
```
The points are one of two colors and the coloring is split along a diagonal line. 

Fourth, we make a scatterplot that color points using the numeric variable. Then we comment on the color scale. 
```{r}
splot2 <- ggplot(p2_df, aes(x = x, y = y, color = vec_num)) + geom_point()
```
The points are colored on a gradient from dark to light and split along a diagonal ine as in the previous plot. Light points are on the upper/right side and dark points are on the lower/left side of the diagonal line. 

Fifth, we make a scatterplot that color points using the factor variable. Then we comment on the color scale.
```{r}
splot3 <- ggplot(p2_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
```
The points are colored and distributed in the same manner as in the first plot. 

Last, we export our first logical variable scatterplot to the project directory using ggsave.
```{r}
ggsave("hw1_p2_splot1.png", plot = splot1, device = png())
```




