---
title: "Homework 1: Amelia Grant-Alfieri, ag3911"
output: github_document
---

# embed R code: 

```{r cars}
summary(cars)
```

# include plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r setup, include = FALSE}
library(tidyverse)
```









The purpose of this file is to present a couple of basic plots using `ggplot`.

First we create a dataframe containing variables for our plots.

# Problem 1

First, we create a dataframe containing four elements:
- A random sample of size 10 from a uniform [0,5] distribution
- A logical vector indicating whether elements of the sample are greater than 2
- A (length-10) character vector
- A (length-10) factor vector. 

```{r df_create}
set.seed(1234)

plot_df = tibble(
  x = rnorm(1000, sd = .5),
  y = 1 + 2 * x + rnorm(1000)
)
```

Second, we caluclate the mean of each variable. Then we analyze what works and what doesn't work and why.

Third, we apply the as.numeric function to the logical, character, and factor variables. Then we describe what happens.

Fourth, we convert the character variable to factor to numeric and convert the factor variable from factor to character to numeric. Then we describe what happens. 

# Problem 2

First, we create a data frame containing five elements:
- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the  x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector-

Second, we describe the vector using inline R code, including: i) the size of the dataset, ii) the mean and median of x, and iii) the proportion of cases for which the logical vector is  TRUE.

Third, we make a scatterplot of y vs x and color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). 

Fourth, we make a scatterplot that color points using the numeric variable. Then we comment on the color scale. 

Fifth, we make a scatterplot that color points using the factor variable. Then we comment on the color scale.

Last, we export our first logical variable scatterplot to the project directory using  ggsave.







First we show a histogram of the `x` variable.

```{r x_hist}
ggplot(plot_df, aes(x = x)) + geom_histogram()
```

Next we show a scatterplot of `y` vs `x`. 

```{r yx_scatter}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

